---
output: pdf_document
title: "A Random Forest on the Titanic"
author: "Antonio Skilton"
date: "September 12, 2016"
---

## The Data
```{r,echo=FALSE,message=FALSE}
library(knitr)
```
```{r,echo=TRUE}
data <- read.csv("train.csv")
survived_label <- ifelse(data$Survived==1,"survived","died")
data$Cabin[data$Cabin==""] <- NA
source("cleanAndManipulate.R")
```
```{r,echo=FALSE}
data2 <- data
data2$Name <- paste(substr(data$Name, 1, 10),"...",sep="")
```

```{r,echo=FALSE}
kable(t(data2[1:5,]),caption="First 5 Observations in Passenger Data",align='c')
```

\newpage

##Missing Data
```{r, message=FALSE, warning=FALSE,echo=TRUE, include=TRUE}
library(VIM)
aggr(data, col=c('navyblue','red'), numbers=TRUE,
     labels=names(data), sortVars=FALSE, cex.axis=.7,
     prop=TRUE, gap=3, ylab=c("Histogram of missing data","Pattern"))
```

\newpage

## Name/Title
#### As seen above, passenger names all have some title associated with them. I suspect that there is infmation to be gleaned from titles in the names that might not be available in either Sex or Age.
```{r}
data$title <- NA
for(string in c("master","mr","miss","mrs")){ 
  data$title[grepl(string,data$Name,ignore.case=T)] <- string}
```

#### After creating a new column for titles, I fill in missing variables. I assume that those missing a traditional title are adults and supply them a replacement title based upon their age and sex.
```{r}
data$title[is.na(data$title)] <- ifelse(data$Sex[is.na(data$title)]=="female","mrs","mr")
```

```{r,echo=FALSE}
data$title <- as.factor(data$title)
kable(data[1:10,c("title","Name","Sex")],caption="The new 'title' column, made from 'Name' and 'Sex'")
```

#### There are clear differences in survival rates that are associated with title. 

```{r, message=FALSE, warning=FALSE, echo=FALSE}
kable(table(survived_label,data$title))
```

\newpage 

##Other variables

###Child

#### Anybody who is either below or equal to the age of 10, below or equal to the age of 10 *and* is accompanied by a parent, is traveling with more than one sibling, or has the name "master" is recorded as a child.
```{r}
data$child <- ifelse(
  data$Age<=10 | (data$Age <= 15 & data$Parch>0) | data$SibSp > 1 | data$title=="master",
  1,0)
```

###Missing Age
####Variable is equal to 1 where passenger has no recorded age.
```{r}
data$ageMissing <- ifelse(is.na(data$Age),1,0)
```

###Cabin
####Variable is equal to 1 where passenger has no recorded cabin.
```{r}
data$cabin <- ifelse(is.na(data$Cabin),1,0)
```

###Ticket number
####The ticket numbers include both special characters and non-numeric characters before I clean them. I then seperate them into groups based upon the uneven distribution of the ticket numbers.
```{r}
data$Ticket_numeric <- as.numeric(gsub("[^0-9]","",data$Ticket))
data$Ticket_group <- as.factor(ifelse(data$Ticket_numeric < 300000,"lowTicketNumber","highTicketNumber"))
```

####Family Size
```{r}
data$familySize <- data$Parch + data$SibSp
```


###Dummy Variables
```{r,echo=FALSE,message=FALSE}
library(caret)
data <- data[data$Embarked!="",]
data$Embarked <- droplevels((data$Embarked))
data$Pclass <- as.factor(data$Pclass)
```

####with the following code we produce the dummy variables for the Embarked, Pclass, title, and Ticket_group columns.
```{r}
data <- cbind(data,
              predict(
                dummyVars(~Embarked + Pclass + title + Ticket_group, data = data),
                newdata = data)
              )
```

\newpage

##Final Edits
####Using the dplyr package I remove columns that through trial and error I have found increase the classification error rate in the random forest model.
```{r, message=FALSE}
library(dplyr)
data<-data %>%
    dplyr::select(-Ticket_numeric,-Sex,-title.master,
                  -title.miss,-Cabin,-Pclass,-title) %>%
    dplyr::select(Survived,child,contains("title"),
                  contains("Pclass"),cabin,contains("group"),
                  Fare,Parch,familySize,ageMissing) 
```

```{r,echo=FALSE}
data <- data%>%
    na.omit() %>%
    mutate_if(function(col) ifelse(is.factor(col),FALSE,max(col)==1 & min(col)==0),as.factor)
```

```{r}
Survived <- select(data, Survived) %>% unlist()  %>% as.factor()
predictors <- select(data, -Survived)
```

\newpage

##Random Forest Model
```{r,message=FALSE}
library(randomForest)
set.seed(1)
randomForest(predictors,Survived,importance=T,mtry=3,xtest=)
```

####We can now test the model on test.csv. The above cleaning and manipulation is performed by the cleanAndManipulate() function.
```{r}
test <- read.csv("test.csv")
source("cleanAndManipulate.r")
test2 <- cleanAndManipulate(test,test=TRUE)
library(randomForest)
set.seed(1)
randomForest(predictors,Survived,importance=T,mtry=3,xtest=test2)
```

####The test error rate os close to the error rate from the training data! Success!
